SPECS-decadal-MOS

### 2017-08-01 SPECS/specs-decade-mos.Rmd
EU-Circle case 1

library(esd)
#library(knitr)
#purl('git/esd_Rmarkdown/SPECS/specs-decade-mos.Rmd', output='git/esd_Rmarkdown/SPECS/specs-decade-mos.R')
#library(ncdf)
setwd("~/R")

downloadfromunican <- FALSE    # Preparations - get the data
downscaledreanalysis <- FALSE  # SM: Assessment of large-small scale link
looexperiments <- FALSE        # Cross-validation MOS on forecasts
diagnoseFCST <- FALSE          # Show correlation & common EOFs of reanalysis+fcsts

@@

## Function to extract a subset and organise the predictor data
## for all ensemble members
Xpredictor <- function(x,m=1:3,it=NULL,is=NULL,mask=FALSE,
                       FUN='mean',verbose=FALSE) {
  if (verbose) print('Xpredictor')
  ## x is the data, m is the forecast interval (years), and it
  ## specifies e.g. months
  ## Get the dimensions of the data structure
  d <- dim(x$Data)
  if (verbose) {print(summary(x)); print(d)}
  ## initialisations members times nx ny

  ## Prepare one member per loop
  for (im in 1:d[2]) {
    index <- x$Dates$start; nt <- length(index); jj <- 1
    for (ii in m) {
      if (verbose) print(paste('member',im,'initialisation',ii))
      ## Read the entire field
      y1 <- fcst2field(x,it0=10-ii,im=im,verbose=verbose)
      ## Extract a sub-region
      y1 <- subset(y1,is=is)
      ## If it!=NULL select subset of the year 
      y1 <- aggregate(subset(y1,it=it),year,FUN)
      ## If mask, remove the land component
      if (mask) y1 <- mask(y1,land=TRUE)
      if (verbose) {print(index(y1)); print(index(lag(zoo(y1),k=-ii)))}
      ## Let Y1 contain the end results: sum the different lead times
      if (jj==1)  Y1 <- zoo(y1) else
                  Y1 <- zoo(Y1) + zoo(lag(y1,k=-ii))
      jj <- jj + 1
    }
    ## Divide by the number of initialisations to get the forecast interval mean
    coredata(Y1) <- coredata(Y1)/length(m)
    ## Set the appropriate attributes and class
    if (verbose) {print('Set the appropriate attributes'); print(names(attributes(y1)))}
    Y1 <- attrcp(y1,Y1)
    if (verbose) print(paste('Set the appropriate class',class(y1),collapse='-'))
    class(Y1) <- class(y1)
    if (verbose) print(class(y1))

@@

  ##Disse går å kombinere til en rad:
  ##  print(paste('cdo ifthen -gtc,0',ifile,ifile,fname))
  ##  system(paste('cdo ifthen -gtc,0',ifile,ifile,fname))
  X <- aggregatedailyfield(fname=fname,cline=cline)
  Y <- aggregatedailyfield(fname=,'mask.nc',cline=cline,output='fw_0.25deg_reg.nc')
  file.remove(fname); file.remove('mask.nc')
}

@@

  if (mask) Z <- mask(Y,land=TRUE) else Z <- Y
  Z <- subset(Z,it=range(year(X)))
  for (i in 1:10) {
    z <- subset(X,it=month.abb[i])
    index(z) <- year(z)
    Z <- combine(Z,z)
  }
  ceof <- EOF(Z)

@@

```{r diagnose}
diagnoseMOS <- function(results,y,l=1) {
  print(paste('diagnoseMOS',results))
  load(results)  
  eval(parse(text=paste('zmos <-',substr(results,14,nchar(results)-4))))
  attr(zmos,'variable') <- varid(y)
  print(range(year(zmos)))
  y <- pentad(y,it0=min(year(zmos)),l=l)
  print(index(y))
  print(index(zmos))
  print('ensemble mean: aggregate(.)')
  ## Different ensemble emmbers are represented as calendar months
  Z <- aggregate(zmos,year,'mean')
  index(Z) <- year(Z)
  print('correlation: ensemble mean FCST & reanalysis')
  r <- corfield(Z,y,plot=FALSE)
  map(r,colbar=list(breaks=seq(-1,1,by=0.1),pal='t2m',rev=TRUE),type='fill',new=FALSE)
  figlab('EOBS+ensemble mean MOS',ypos=0.999)

@@

print('Get the predictands: mu & fw from EOBS')
if (!file.exists('~/data/mu_0.25deg_reg.nc'))
  mu.eobs <-  wetdaymean('~/data/data.ECAD/rr_0.25deg_reg.nc') else
  mu.eobs <- retrieve('~/data/mu_0.25deg_reg.nc')
fw.eobs <- retrieve('~/data/fw_0.25deg_reg.nc')
attr(mu.eobs,'variable') <- 'mu'
attr(fw.eobs,'variable') <- 'fw'
attr(fw.eobs,'unit') <- 'fraction'
index(fw.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
index(mu.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))

```{r ds}

@@

## Also get the mean sea-level pressure for the wet-day frequency 
  slp <- annual(retrieve('slp.mon.mean.nc',lon=c(-60,40),lat=c(20,70)))
  slp.eof <- EOF(slp)

@@

print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
load('~/data/gfdl_DECA.rda')

Leave-one-out cross-validation experiments

```{r loo}

if (looexperiments) {
## Carry out a LOO cross-validation experiment
  print('Check results for LOO experiment - if absent do it')
  if (!file.exists('specsdecadal.z.fw.y1.rda')) {
    print('do specsdecadal.z.fw.y1.rda')
    z.fw.y1 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1,it=NULL,is=NULL,mask=FALSE)
    save(file='specsdecadal.z.fw.y1.rda',z.fw.y1)
    rm("z.fw.y1"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.mu.y1.rda')) {
    print('do specsdecadal.z.mu.y1.rda')
    z.mu.y1 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.rda',z.mu.y1)
    rm("z.mu.y1"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.fw.y1.3.rda')) {
    print('do specsdecadal.z.fw.y1.3.rda')
    z.fw.y1.3 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:3,it=NULL,is=NULL,mask=FALSE)
    save(file='specsdecadal.z.fw.y1.3.rda',z.fw.y1.3)
    rm("z.fw.y1.3"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.mu.y1.3.rda')) {
    print('do specsdecadal.z.mu.y1.3.rda')
    z.mu.y1.3 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:3,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.3.rda',z.mu.y1.3)
    rm("z.mu.y1.3"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.fw.y1.5.rda')) {
    print('do specsdecadal.z.fw.y1.5.rda')
    z.fw.y1.5 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:5,it=NULL,is=NULL,mask=FALSE)
    save(file='specsdecadal.z.fw.y1.5.rda',z.fw.y1.5)
    rm("z.fw.y1.5"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.mu.y1.5.rda')) {
    print('do specsdecadal.z.mu.y1.5.rda')
    z.mu.y1.5 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:5,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.5.rda',z.mu.y1.5)
    rm("z.mu.y1.5"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.fw.y1.9.rda')) {
    print('do specsdecadal.z.fw.y1.9.rda')
     z.fw.y1.9 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:9,it=NULL,is=NULL,mask=FALSE)
     save(file='specsdecadal.z.fw.y1.9.rda',z.fw.y1.9)
     rm("z.fw.y1.9"); gc(reset=TRUE)
   }
  if (!file.exists('specsdecadal.z.mu.y1.9.rda')) {
    print('do specsdecadal.z.mu.y1.9.rda')
    z.mu.y1.9 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:9,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.9.rda',z.mu.y1.9)
    rm("z.mu.y1.9"); gc(reset=TRUE)
  }

## Diagnostics:
  print('diagnose the MOS results')
  load('specsdecadal.z.mu.y1.3.rda')
  it0 <- range(year(z.mu.y1.3))
  print(paste('time period',paste(it0,collapse='-')))
  score.mu.y1 <- diagnoseMOS('specsdecadal.z.mu.y1.rda',mu.eobs)
  score.fw.y1 <- diagnoseMOS('specsdecadal.z.fw.y1.rda',fw.eobs)
  score.mu.y1.3 <- diagnoseMOS('specsdecadal.z.mu.y1.3.rda',mu.eobs,l=3)
  score.fw.y1.3 <- diagnoseMOS('specsdecadal.z.fw.y1.3.rda',fw.eobs,l=3)
  score.mu.y1.5 <- diagnoseMOS('specsdecadal.z.mu.y1.5.rda',mu.eobs,l=5)
  score.fw.y1.5 <- diagnoseMOS('specsdecadal.z.fw.y1.5.rda',fw.eobs,l=5)
  score.mu.y1.9 <- diagnoseMOS('specsdecadal.z.mu.y1.9.rda',mu.eobs,l=9)
  score.fw.y1.9 <- diagnoseMOS('specsdecadal.z.fw.y1.9.rda',fw.eobs,l=9)
  while (dev.cur()>1) dev.off()
## Repeat and make a table of the results...

}

Diagnose the forecasts

```{r diagnosefcst}
## Common EOFs to diagnose the results?

if (diagnoseFCST) {
  comparemannual(tas_DECA,mask=TRUE,l= 1)
  comparemannual(tas_DECA,mask=TRUE,l= 2)
  comparemannual(tas_DECA,mask=TRUE,l= 4)
  comparemannual(tas_DECA,mask=TRUE,l= 5)
  comparemannual(tas_DECA,mask=TRUE,l= 6)
  comparemannual(tas_DECA,mask=TRUE,l= 7)
  comparemannual(tas_DECA,mask=TRUE,l= 8)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 1)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 2)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 3)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 4)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 5)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 6)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 7)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 8)
  rmap.1.10.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3, eofs=1:10)
  rmap.1.10.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:10)
}
```

```{r rmaps}
  ## Test the forecasts of the larg-scale anomalies/gravest EOFs
  rmap.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3)
  rmap.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3)
  rmap.1.3.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3, eofs=1:3)
  rmap.1.3.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:3)
  rmap.1.5.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:5)
  rmap.1.2.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:2)
  rmap.1.9.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:9)
```

```{r rdiff}
  rdiff <- rmap.1.3.nomask.3 - rmap.nomask.3
  map(rdiff,new=FALSE)
  figlab('TAS forecast (3 years): correlations EOFs 1-3 vs full field',ypos=0.99)

@@

  rdiff <- rmap.1.9.nomask.3 - rmap.nomask.3
  map(rdiff,new=FALSE)
  figlab('TAS forecast (3 years): correlations EOFs 1-9 vs full field',ypos=0.99)
```

@@

```{r slp}

## Repeat for SLP
slp <- retrieve('~/data/ncep/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
slp <- anomaly(slp)
yst <- scaledependency(slp)
image(attr(yst,'timescales'),attr(yst,'gridboxes'),t(t(yst)/colMeans(yst)),
      main='Mean sea-level pressure: spectral power',
      xlab='Temporal scale (years)',ylab='spatial scale (grid boxes)')
contour(attr(yst,'timescales'),attr(yst,'gridboxes'),yst,add=TRUE)

>>>>

library(esd)
#library(knitr)
#purl('git/esd_Rmarkdown/SPECS/specs-decade-mos.Rmd', output='git/esd_Rmarkdown/SPECS/specs-decade-mos.R')
#library(ncdf)
setwd("~/R")
experiments <- c('~/data/gfdl_DECA.rda','~/data/hadcm3_DECA.rda','~/data/miroc5_DECA.rda','~/data/mpi_DECA.rda')

downloadfromunican <- FALSE    # Preparations - get the data
downscaledreanalysis <- TRUE  # SM: Assessment of large-small scale link
looexperiments <- TRUE        # Cross-validation MOS on forecasts
diagnoseFCST <- TRUE          # Show correlation & common EOFs of reanalysis+fcsts

@@

## Function to extract a subset and organise the predictor data
## for all ensemble members
Xpredictor <- function(x,m=1:3,it=NULL,is=NULL,mask=FALSE,
                       FUN='mean',verbose=FALSE) {
  if (verbose) print('Xpredictor')
  ## x is the data, m is the forecast interval (years), and it
  ## specifies e.g. months
  ## Get the dimensions of the data structure
  d <- dim(x$Data)
  nmembers <- length(x$members)
  if (verbose) {print(summary(x)); print(d)}
  ## initialisations members times nx ny

  ## Prepare one member per loop
  for (im in 1:d[2]) {
    index <- x$Dates$start; nt <- length(index); jj <- 1
    for (ii in m) {
      if (verbose) print(paste('member',im,'initialisation',ii))
      ## Read the entire field
      y1 <- fcst2field(x,it0=nmembers-ii,im=im,verbose=verbose)
      ## Extract a sub-region
      y1 <- subset(y1,is=is)
      ## If it!=NULL select subset of the year 
      y1 <- aggregate(subset(y1,it=it),year,FUN)
      ## If mask, remove the land component
      if (mask) y1 <- mask(y1,land=TRUE)
      if (verbose) {print(index(y1)); print(index(lag(zoo(y1),k=-ii)))}
      ## Let Y1 contain the end results: sum the different lead times
      if (jj==1)  Y1 <- zoo(y1) else
                  Y1 <- zoo(Y1) + zoo(lag(y1,k=-ii))
      jj <- jj + 1
    }
    ## Divide by the number of initialisations to get the forecast interval mean
    coredata(Y1) <- coredata(Y1)/length(m)
    ## Set the appropriate attributes and class
    if (verbose) {print('Set the appropriate attributes'); print(names(attributes(y1)))}
    Y1 <- attrcp(y1,Y1)
    if (verbose) print(paste('Set the appropriate class',class(y1),collapse='-'))
    class(Y1) <- class(y1)
    if (verbose) print(class(y1))

@@

  ##Disse går å kombinere til en rad:
  ##  print(paste('cdo ifthen -gtc,0',ifile,ifile,fname))
  ##  system(paste('cdo ifthen -gtc,0',ifile,ifile,fname))
  X <- aggregatedailyfield(fname=fname,cline=cline)
  Y <- aggregatedailyfield(fname='mask.nc',cline=cline,output='fw_0.25deg_reg.nc')
  file.remove(fname); file.remove('mask.nc')
}

@@

  if (mask) Z <- mask(Y,land=TRUE) else Z <- Y
  Z <- subset(Z,it=range(year(X)))
  nmembers <- length(table(month(X)))
  for (i in 1:nmembers) {
    z <- subset(X,it=month.abb[i])
    index(z) <- year(z)
    Z <- combine(Z,z)
  }
  ceof <- EOF(Z)

@@

```{r diagnose}
diagnoseMOS <- function(results,y,l=1) {
  print(paste('diagnoseMOS',results))
  load(results)  
  zl <- regexpr('.z.',results)
  eval(parse(text=paste('zmos <-',substr(results,zl+1,nchar(results)-4))))
  attr(zmos,'variable') <- varid(y)
  print(range(year(zmos)))
  y <- pentad(y,it0=min(year(zmos)),l=l)
  print(index(y))
  print(index(zmos))
  print('ensemble mean: aggregate(.)')
  ## Different ensemble emmbers are represented as calendar months
  Z <- aggregate(zmos,year,'mean')
  index(Z) <- year(Z)
  print('correlation: ensemble mean FCST & reanalysis')
  r <- corfield(Z,y,plot=FALSE)
  map(r,colbar=list(breaks=seq(-1,1,by=0.1),pal='t2m',rev=TRUE),type='fill',new=FALSE)
  figlab('EOBS+ensemble mean MOS',ypos=0.999)

@@

print('Get the predictands: mu & fw from EOBS')
if (!file.exists('~/mu_0.25deg_reg.nc'))
  mu.eobs <-  wetdaymean('~/data/data.ECAD/rr_0.25deg_reg.nc') else
  mu.eobs <- retrieve('~/mu_0.25deg_reg.nc')
fw.eobs <- retrieve('~/fw_0.25deg_reg.nc')
attr(mu.eobs,'variable') <- 'mu'
attr(fw.eobs,'variable') <- 'fw'
attr(fw.eobs,'unit') <- 'fraction'
#index(fw.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
#index(mu.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
```

```{r ds,fig.height=8}

@@

## Also get the mean sea-level pressure for the wet-day frequency 
  slp <- annual(retrieve('~/Downloads/slp.mon.mean.nc',lon=c(-60,40),lat=c(20,70)))
  slp.eof <- EOF(slp)

@@

for (iexp in 1:1) {
  print(paste('Get the decadal predictors saved locally in',experiments[iexp]))
  load(experiments[iexp])

## Leave-one-out cross-validation experiments
## Names of file names with results stored  
fw1name <- paste('specsdecadal.z.fw.y1.gcm',iexp,'rda',sep='.')
mu1name <- paste('specsdecadal.z.mu.y1.gcm',iexp,'rda',sep='.')
fw3name <- paste('specsdecadal.z.fw.y1.3.gcm',iexp,'rda',sep='.')
mu3name <- paste('specsdecadal.z.mu.y1.3.gcm',iexp,'rda',sep='.')
fw5name <- paste('specsdecadal.z.fw.y1.5.gcm',iexp,'rda',sep='.')
mu5name <- paste('specsdecadal.z.mu.y1.5.gcm',iexp,'rda',sep='.')
fw9name <- paste('specsdecadal.z.fw.y1.9.gcm',iexp,'rda',sep='.')
mu9name <- paste('specsdecadal.z.mu.y1.9.gcm',iexp,'rda',sep='.')

if (looexperiments) {
## Carry out a LOO cross-validation experiment
  print('Check results for LOO experiment - if absent do it')
  if (!file.exists(fw1name)) {
    print(fw1name)
    z.fw.y1 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1,it=NULL,is=NULL,mask=FALSE)
    save(file=fw1name,z.fw.y1)
    rm("z.fw.y1"); gc(reset=TRUE)
  }

  if (!file.exists(mu1name)) {
    print(mu1name)
    z.mu.y1 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1,it=NULL,is=NULL,mask=TRUE)
    save(file=mu1name,z.mu.y1)
    rm("z.mu.y1"); gc(reset=TRUE)
  }
  if (!file.exists(fw3name)) {
    print(fw3name)
    z.fw.y1.3 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:3,it=NULL,is=NULL,mask=FALSE)
    save(file=fw3name,z.fw.y1.3)
    rm("z.fw.y1.3"); gc(reset=TRUE)
  }
  if (!file.exists(mu3name)) {
    print(mu3name)
    z.mu.y1.3 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:3,it=NULL,is=NULL,mask=TRUE)
    save(file=mu3name,z.mu.y1.3)
    rm("z.mu.y1.3"); gc(reset=TRUE)
  }
  if (!file.exists(fw5name)) {
    print(fw5name)
    z.fw.y1.5 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:5,it=NULL,is=NULL,mask=FALSE)
    save(file=fw5name,z.fw.y1.5)
    rm("z.fw.y1.5"); gc(reset=TRUE)
  }
  if (!file.exists(mu5name)) {
    print(mu5name)
    z.mu.y1.5 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:5,it=NULL,is=NULL,mask=TRUE)
    save(file=mu5name,z.mu.y1.5)
    rm("z.mu.y1.5"); gc(reset=TRUE)
  }
  if (!file.exists(fw9name)) {
    print(fw9name)
     z.fw.y1.9 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:9,it=NULL,is=NULL,mask=FALSE)
     save(file=fw9name,z.fw.y1.9)
     rm("z.fw.y1.9"); gc(reset=TRUE)
   }
  if (!file.exists(mu9name)) {
    print(mu9name)
    z.mu.y1.9 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:9,it=NULL,is=NULL,mask=TRUE)
    save(file=mu9name,z.mu.y1.9)
    rm("z.mu.y1.9"); gc(reset=TRUE)
  }

## Diagnostics:
  print('diagnose the MOS results')
  load('specsdecadal.z.mu.y1.3.rda')
  it0 <- range(year(z.mu.y1.3))
  print(paste('time period',paste(it0,collapse='-')))
  score.mu.y1 <- diagnoseMOS('specsdecadal.z.mu.y1.rda',mu.eobs)
  score.fw.y1 <- diagnoseMOS('specsdecadal.z.fw.y1.rda',fw.eobs)
  score.mu.y1.3 <- diagnoseMOS('specsdecadal.z.mu.y1.3.rda',mu.eobs,l=3)
  score.fw.y1.3 <- diagnoseMOS('specsdecadal.z.fw.y1.3.rda',fw.eobs,l=3)
  score.mu.y1.5 <- diagnoseMOS('specsdecadal.z.mu.y1.5.rda',mu.eobs,l=5)
  score.fw.y1.5 <- diagnoseMOS('specsdecadal.z.fw.y1.5.rda',fw.eobs,l=5)
  score.mu.y1.9 <- diagnoseMOS('specsdecadal.z.mu.y1.9.rda',mu.eobs,l=9)
  score.fw.y1.9 <- diagnoseMOS('specsdecadal.z.fw.y1.9.rda',fw.eobs,l=9)

## Repeat and make a table of the results...
}

## Diagnose the forecasts

## Common EOFs to diagnose the results?

if (diagnoseFCST) {
  comparemannual(tas_DECA,mask=TRUE,l= 1)
  comparemannual(tas_DECA,mask=TRUE,l= 2)
  comparemannual(tas_DECA,mask=TRUE,l= 4)
  comparemannual(tas_DECA,mask=TRUE,l= 5)
  comparemannual(tas_DECA,mask=TRUE,l= 6)
  comparemannual(tas_DECA,mask=TRUE,l= 7)
  comparemannual(tas_DECA,mask=TRUE,l= 8)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 1)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 2)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 3)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 4)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 5)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 6)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 7)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 8)
  rmap.1.10.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3, eofs=1:10)
  rmap.1.10.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:10)
}

  ## Test the forecasts of the larg-scale anomalies/gravest EOFs
  rmap.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3)
  rmap.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3)
  rmap.1.3.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3, eofs=1:3)
  rmap.1.3.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:3)
  rmap.1.5.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:5)
  rmap.1.2.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:2)
  rmap.1.9.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:9)

  rdiff <- rmap.1.3.nomask.3 - rmap.nomask.3
  map(rdiff,new=FALSE)
  figlab('TAS forecast (3 years): correlations EOFs 1-3 vs full field',ypos=0.99)

@@

  rdiff <- rmap.1.9.nomask.3 - rmap.nomask.3
  map(rdiff,new=FALSE)
  figlab('TAS forecast (3 years): correlations EOFs 1-9 vs full field',ypos=0.99)
}
```

@@

```{r slp}
## Repeat for SLP
slp <- retrieve('~/data/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
slp <- anomaly(slp)
yst <- scaledependency(slp)
image(attr(yst,'timescales'),attr(yst,'gridboxes'),t(t(yst)/colMeans(yst)),
      main='Mean sea-level pressure: spectral power',
      xlab='Temporal scale (years)',ylab='spatial scale (grid boxes)')
contour(attr(yst,'timescales'),attr(yst,'gridboxes'),yst,add=TRUE)
